Adapting the Chumbley Score to Bullet Striations

### 2018-04-23 presentation/presentation.Rmd
presentation

---
title: "Adaption of the Chumbley Score to matching of bullet striation marks"
authors:
- affiliation: Department of Statistics, Iowa State University 
  name: Ganesh Krishnan
  thanks: The authors gratefully acknowledge ...
- affiliation: Department of Statistics and CSAFE, Iowa State University 
  name: Heike Hofmann
biblio-style: apsr
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    template: template.tex
  html_document: default
blinded: 0
keywords:
- 3 to 6 keywords
- that do not appear in the title
bibliography: bibliography
abstract: null
---

\newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
\newcommand{\gk}[1]{{\textcolor{green}{#1}}}
\newcommand{\cited}[1]{{\textcolor{red}{#1}}}
\setlength\parindent{0pt}

\tableofcontents
\newpage
```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  collapse = TRUE,
  comment = "",
  fig.height = 6,
  fig.width = 6,
  fig.align = "center",
  out.width= '\\textwidth',
  cache = FALSE,
  fig.path='figures/',
  echo=FALSE,
  cache=TRUE
)
options(knitr.table.format = "latex")
library(tidyverse)
library(kableExtra)
library(xtable)
library(gridExtra)
```

@@

```{r data}
if (!file.exists("../data/all-sigs.rds")) {
  files <- dir("../data/signatures", pattern="csv")
  all <- data.frame()
  for (file in files) {
    tmp <- read.csv(file=file.path(path="../data/signatures", file)) 
    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% dplyr::select(-x)
    all <- rbind(all, tmp)
  }
  all <- all %>% filter(land1_id < land2_id)
  saveRDS(all, file="../data/all-sigs.rds")
} else {
  all <- readRDS("../data/all-sigs.rds")
}

@@

```{r data-profiles}
if (!file.exists("../data/all-profiles.rds")) {
  files <- dir("../data/profiles/", pattern="csv")
  allp <- data.frame()
  for (file in files) {
    tmp <- read.csv(file=file.path(path="../data/profiles", file)) 
    #print(names(tmp))
    if (length(grep("coarse",names(tmp))) == 0)
      tmp$coarse <- gsub(".*coar-(.*).csv", "\\1", file)
    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% dplyr::select(-x)
    allp <- rbind(allp, tmp)
  }
  allp <- allp %>% filter(land1_id < land2_id)
  allp <- allp %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
  saveRDS(allp, file="../data/all-profiles.rds")
} else {
  allp <- readRDS("../data/all-profiles.rds")
}

@@

if (!file.exists("../data/invprofiles.rds")) {
  files <- dir("../data/inverse-profiles/", pattern="rds")
  allinv <- data.frame()
  for (file in files) {
    tmp <- readRDS(file=file.path(path="../data/inverse-profiles", file)) 
    allinv <- rbind(allinv, tmp)
  }
  allinv <- allinv %>% filter(land1_id < land2_id)
  allinv <- allinv %>% dplyr::select(-coarseness)
  saveRDS(allinv, file="../data/invprofiles.rds")
} else {
  allinv <- readRDS("../data/invprofiles.rds")
}

@@

```{r data-sig-coars}
if (!file.exists("../data/all-sig-c.rds")) {
  files <- dir("../data/sig_diff_coarseness/", pattern="csv")
  all_sig_c <- data.frame()
  for (file in files) {
    tmp <- read.csv(file=file.path(path="../data/sig_diff_coarseness/", file)) 
    #print(names(tmp))
    if (length(grep("coarse",names(tmp))) == 0)
      tmp$coarse <- gsub(".*coarse-(.*).csv", "\\1", file)
    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% dplyr::select(-x)
    all_sig_c <- rbind(all_sig_c, tmp)
  }
  all_sig_c <- all_sig_c %>% filter(land1_id < land2_id)
  all_sig_c <- all_sig_c %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
  saveRDS(all_sig_c, file="../data/all-sig-c.rds")
} else {
  all_sig_c <- readRDS("../data/all-sig-c.rds")
}

@@

# Introduction and Background

## Motivation
Same source analyses are a major part of an Forensic Toolmark Examiner's job. In current practice examiners  make these comparisons by visual inspection under a comparison microscope and come to one of the following four conclusions: identification, inconclusive, elimination or unsuitable for examination~\citep{afte-toolmarks1998}. These conclusions are made on the basis of "unique surface contours" of the two toolmarks being in "sufficient agreement" \citep{afte-toolmarks1998}. AFTE describes the term "sufficient agreement" as the possibility of another tool producing the markings under comparison, as practically impossible \citep{afte-toolmarks1998}. Potential subject bias in the assessment as well as the lack of specified error rates are the main points of criticisms  first raised by the National Research Council in 2009 \citep{NAS:2009} and later emphasized further by the President's Council of Advisors on Science and Technology \citep{pcast2016}.

Technological advances, such as profilometers and confocal microscopy allow to measure 3D surfaces in a high-resolution digitized form. This technology has become more accessible over the last decade, and has made its way into topological images of ballistics evidence, such as bullet lands and breech faces \citep{DeKinder1, DeKinder2, Bachrach1, vorburger2016}.
Digitized images of 3D surfaces of form the data basis of statistical analysis of toolmarks. A statistical approach based on data  removes both subjectivity from the assessment and allows a quantification of error rates for both false positive and false negative identifications.

\hh{In the next page and a half it is easy to lose the red line. It might help to include a table with an overview.
The table should include the reference to the paper, the data used, the statistical method and the associated error rates.}
Various toolmarks have been studied in the literature:  \citet{manytoolmarks1} and \citet{chumbley} have been analyzing screwdriver marks  digitized using a profilometer;  \citet{manytoolmarks2} have investigated 3D marks from screwdriver, tongue and groove pliers captured using a confocal microscope; \citet{afte-chumbley} have been investigated digitized marks from slip-joint pliers generated by a surface profilometer.

\hh{We need an additional sentence here to get from the data to the statistical methods ... }
\citet{manytoolmarks2} define a relative distance metric and use it as similarity measure between two toolmarks. \citet{manytoolmarks1} extract many small segments in the markings of two toolmarks and compare similiarity using a maximum pearson correlation coefficient. 
The Chumbley scoring method, first introduced by \citet{chumbley}, uses a similar but more extensive framework based on a Mann-Whitney U test of the resulting correlation coefficients. This approach is non-deterministic, because segments are chosen randomly.  \citep{hadler} make the score deterministic for each pair of toolmarks by choosing segments for comparison systematically. This approach also ensures independence between segements of striae.  In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler} to assess striation marks on bullet lands for same-source identification.

Striation marks on bullets are made by impurities in the barrel. As the bullet travels through the barrel, these imperfections leave "scratches" on the bullet surface. Typically, only striation marks in the land engraved areas (LEAs) are considered \citet{afte-article1992}. Bullet lands are depressed areas between the  grooves made by the rifling action of the barrel. Compared to toolmarks made by screwdrivers striation marks on bullets are typically much smaller, both in length and in width. Bullets also have a curved cross-sectional topography. Figure \ref{fig:rgl} shows us how the signature from a bullet land (bottom) lines up with the image of the land (top) from which it was extracted. We can also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.

Bullet matching methods are usually based on these associated signatures. \citet{chu2013} use an automatic method for counting consecutive matching striae (CMS). The authors report an error rate of 52% of the known same source lands comparisons as misidentified (false negative) and zero false positives for known different source lands. \citet{ma2004} and \citet{vorburger2011} discuss CCF (cross-correlation function) and its discriminating power and applicability for same-source analyses of bullets, but do not provide any error rates in their discussion. \citet{aoas} use multiple features like CCF, CMS, D (distance measure) etc in a random forest based method and compare every land against every other land of  digitised versions of Hamby 252 and Hamby 44 \citep{hamby} published on the NIST Ballistics Database \citep{nist}. The authors report an out-of-bag overall error rate of 0.46%, comprised of an error rate of 30.05% of  same-source pairs that were not identified and an error rate of 0.026% of different-source pairs that were incorrectly identified as same-source.

The Chumbley score  provides us with another approach in the same-source assessment of bullet striation marks. \citet{chumbley} compare two toolmarks for same-source. The data for this study was obtained from 50 sequentially manufactured screwdriver tips. \citet{chumbley} report error rates for markings made by the tips at different angles. For markings made at 30 degree the authors  report an average false negative error rate of 0.089  and an average false positive error rate of 0.023. For other angles of 60 and 85 degrees the false negatives error rate is 0.09 while the rate of false positives decreases to 0.01. The paper by \citet{hadler} is based on the same data but the authors focus on  markings made under the same angle. The error rates associated with the deterministic version of the score are  0.06 for false  negatives and a false positive error rate of 0.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.png}
```{r, eval=FALSE, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
land <- read.csv("../data/b6b2l6.csv")
land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw() +
  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
  xlab(expression(paste("Relative Location (in ",mu,m,")", sep="")))
```

\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification (top) and a chart of the corresponding signature of the same land (bottom). The dotted lines connect some peaks visible in both visualizations.}

\end{figure}

## Scans for land engraved areas

Comparisons of striae from bullets are usually based on comparisons of striae in land engraved areas, which are extracted in form of cross sections, called *profiles* \citep{aoas,ma2004}. From profiles bullet *signatures* \citep{chu2013,aoas} are extracted as residuals of a loess fit or Gaussian filter. This effectively removes topographic structure from the data in the attempt to increase the signal to noise ratio. The span of the loess fit was found using cross-validation, as described by \citet{aoas}.

\hh{don't split the discussion on the size. between the next paragraph and }
<!--The majority of Bullet profiles and signatures of the Hamby study \citep{hamby} used in this paper are extracted by procedures mentioned by  \citet{aoas}.--> 
There are two sources of scans for sets from the Hamby study available to us: scans of Hamby 44 and Hamby 252 are available from the NIST database \citep{nist}. Hamby 44 has also been made available to us and has been scanned locally for CSAFE at the Roy J.\ Carver High Resolution Microscopy Facility using a Sensofar confocal light microscope. 
Scans in the NIST database are made with a NanoFocus at 20x magnification. The resolutions of the two instruments are different: the NIST scans are taken at a resolution of 1.5625 $\mu m$ per pixel, while the CSAFE scans are available at a resolution of  0.645 $\mu m$ per pixel.
The length of an average bullet land from Hamby (9 mm Ruger P85) is about 2 millimeter, resulting in signatures of about 1200 pixels for NIST scans, and about 3000 pixels for CSAFE scans.

In comparison, scans from the profilometer used by \citet{chumbley, hadler} were taken at a resolution of about 0.73 $\mu m$ per pixel. The screw driver toolmarks are  about 7 mm in length \citep{manytoolmarks1}, for a total of over 9000 pixels for the width of these scans.

This severe limitation in the amount of available data poses the main challenge in adapting the Chumbley score to matching bullet lands, because of the resulting loss in power.

## The Chumbley Score Test

The Chumbley score algorithm takes input in form of two digitized toolmarks. The toolmark is in form of $z(t)$ which is a spatial process for location indexed by $t$. $t$ here denotes equally spaced pixel locations for the striation marks under consideration, $t = 1, ..., T$. Let further $z^s(t)$ denote a vector of markings of length $s$ starting in location  $t$.

Let $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$ be two digitized toolmarks (where $T_1$ and $T_2$ are not necessarily equal). 
<!--The vectorized processes representing two sets of striaes are therefore shown as $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$. Here $x$ and $y$ denote two instances of the process $z(*)$ which means two striation marks, while the indexing $t$ or pixel range is shown as $t_1$ and $t_2$ for $x$ and $y$ respectively. This representation of $t$ as $t_1$ and $t_2$ lets us identify them as either of same or different lengths.--> 
The toolmarks under consideration are potentially from two different sources or the same source. $T_1$ and $T_2$, as represented above, are the final pixel indexes of each marking and therefore give the respective lengths of the markings.

In a pre-processing step  the two markings are smoothed using a lowess \citep{lowess} with coarseness parameter $c$. 
 Originally, this smoothing is intended to remove drift and (sub)class characteristics from individual markings, however, in the setting of matching bullet striae, we can also make use of this mechanism to separate bullet curvature in profiles from signatures before matching signatures.

After removing sub-class structure, 
the Chumbley scores is calculated in two steps: an optimization step and a validation step.
In the optimization step, the two markings are aligned horizontally such that within a pre-defined window of length $w_o$ the correlation between $x(t_1)$ and $y(t_2)$ is maximized:
\[
\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x^{w_o} (t_1), y^{w_o}(t_2) \right)
\]
This results in an optimal vertical (in-phase) shift of $t_1^o - t_2^o$ for aligning the two markings. \hh{We will denote the relative optimal locations as $t_1^*$ and $t_2^*$, where $t_i^* = t_i^o/(T_i-w_o)$ for $i=1,2$, such that $t_1^*, t_2^* \in [0,1]$. Once (sub-)class characteristics are removed, the relative optimal locations should be distributed according to a uniform distribution in $[0,1]$. }

In the validation step,  two sets of windows of size $w_v$ are chosen from both markings (see Figure \ref{fig:win-comparison}). In the first set, pairs of windows are extracted from the two markings using the optimal vertical shift as determined in the first step, whereas for the second set the windows are extracted using a different (out-of-phase) shift.

More precisely, let us define starting points $s_i^{(k)}$ for each signature $k = 1, 2$ as 
\begin{eqnarray}\label{eq.start}
s^{(k)}_i = 
\begin{cases}
t_k^* + i w_v & \text{ for } i < 0 \\
t_k^* + w_ o + i w_v & \text{ for } i \ge 0,
\end{cases}
\end{eqnarray} for integer values of $i$ with $0 <  s^{(k)}_i \le T_k - w_v$.

Same-shift pairs of length $w_v$ are defined in \citet{hadler} as all 
pairs $(s_i^{(1)}, s_i^{(2)})$ for all integer values $i$ for which both $s_i^{(1)}$ and  $s_i^{(2)}$ are defined. Similarly, different-shift pairs are defined as $(s_i^{(1)}, s_{-i-1}^{(2)})$ for all $i$ where both $s_i^{(1)}$ and  $s_{-i-1}^{(2)}$ are defined (see \autoref{sketch-same-diff}).

\begin{figure}[hbtp]
\centering
\includegraphics[width=.7\textwidth]{images/sketch-same.png}

\includegraphics[width=.7\textwidth]{images/sketch-diff.png}
\caption{\label{sketch-same-diff}Sketch of same-shift pairings  (top) and different-shift pairings (bottom). Filled in rectangles show pairings resulting in correlations, unfilled rectangles are segments without a match.}
\end{figure}

@@

```{r win-comparison-2, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair.", eval=FALSE}

asd<- all %>% filter(wo==120, wv==50, match == TRUE, land1_id == 1)

sigs_graphics<- readRDS("../data/sigs_generate_window.rds")

# Matching signature
#gi<-
#sigs_graphics.match<- sigs_graphics.match %>% filter(land_id == c(1,9))  
#sigs_graphics.match<- sigs_graphics.match %>% select(y, l30, land_id)
d1<- sigs_graphics.match[which(sigs_graphics.match$land_id==1,arr.ind = TRUE),]
d2<- sigs_graphics.match[which(sigs_graphics.match$land_id==19,arr.ind = TRUE),]
data1<-d1 %>%dplyr::select(l30)
#data1<- data1$y
data2<- d2 %>%dplyr::select(l30)
#data2<- data2$y
window_opt = 120 
window_val = 50
coarse = 1
data1<- matrix(unlist(data1))
data2<- matrix(unlist(data2))

  s1.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,2], xmax= max_corr_opt_loc[1,2]+window_opt, ymin=-Inf, ymax=Inf)
    s1.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,2]-window_val, xmax= max_corr_opt_loc[1,2]- 2*window_val, ymin=-Inf, ymax=Inf)
       s1.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,2]+window_opt+ window_val, xmax= max_corr_opt_loc[1,2]+ window_opt+2*window_val, ymin=-Inf, ymax=Inf)

  s2.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,1], xmax= max_corr_opt_loc[1,1]-window_opt, ymin=-Inf, ymax=Inf)
    s2.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,1]-window_opt-window_val, xmax= max_corr_opt_loc[1,1]-window_opt- 2*window_val, ymin=-Inf, ymax=Inf)
       s2.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,1]+ window_val, xmax= max_corr_opt_loc[1,1]+ 2*window_val, ymin=-Inf, ymax=Inf)

p1<-  ggplot(data = d1, aes(x = y, y = l30)) + geom_line() + 
  geom_rect(data=s1.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
    geom_rect(data=s1.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
    geom_rect(data=s1.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
  theme_bw() + labs(y = "Signature Land Id = 1")

 p2<-   ggplot(data = d2, aes(x = y, y = l30)) + geom_line() + 
   geom_rect(data=s2.rect.opt, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "black",alpha=0.5, inherit.aes = FALSE)+
    geom_rect(data=s2.rect.sameshift1, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "red",alpha=0.1, inherit.aes = FALSE)+
    geom_rect(data=s2.rect.sameshift2, aes(xmin=xmin, xmax=xmax , ymin=ymin, ymax=ymax),color = "blue",alpha=0.1, inherit.aes = FALSE)+
  theme_bw() + labs(y = "Signature Land Id = 9")
 ##########

 d1.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,2], xmax= max_corr_opt_loc[1,2]+window_opt, ymin=-Inf, ymax=Inf)
    d1.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,2]-2*window_opt, xmax= max_corr_opt_loc[1,2]-2*window_opt-window_val, ymin=-Inf, ymax=Inf)
       d1.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,2]+window_opt+ window_val, xmax= max_corr_opt_loc[1,2]+ window_opt+2*window_val, ymin=-Inf, ymax=Inf)

  d2.rect.opt <- data.frame(xmin=max_corr_opt_loc[1,1], xmax= max_corr_opt_loc[1,1]-window_opt, ymin=-Inf, ymax=Inf)
    d2.rect.sameshift1 <- data.frame(xmin=max_corr_opt_loc[1,1]-window_opt-window_val, xmax= max_corr_opt_loc[1,1]-window_opt- 2*window_val, ymin=-Inf, ymax=Inf)
       d2.rect.sameshift2 <- data.frame(xmin=max_corr_opt_loc[1,1]+ 4*window_val, xmax= max_corr_opt_loc[1,1]+ 5*window_val, ymin=-Inf, ymax=Inf)

>>>>

---
title: "Adapting the Chumbley Score to Bullet Striations"
author: "Ganesh Krishnan, Heike Hofmann"
date: "April 21, 2018"
header-includes:
   - \usepackage{cleveref}
   - \usepackage{amsmath}
   - \usepackage{amsfonts}
output: 
  beamer_presentation:
    citation_package: natbib
bibliography: bibliography.bib
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  collapse = TRUE,
  comment = "",
  fig.height = 6,
  fig.width = 6,
  fig.align = "center",
  out.width= '\\textwidth',
  cache = FALSE,
  fig.path='figures/',
  echo=FALSE,
  cache=TRUE
)
options(knitr.table.format = "latex")
library(tidyverse)
library(kableExtra)
library(xtable)
library(gridExtra)
```

@@

```{r data}
if (!file.exists("../data/all-sigs.rds")) {
  files <- dir("../data/signatures", pattern="csv")
  all <- data.frame()
  for (file in files) {
    tmp <- read.csv(file=file.path(path="../data/signatures", file)) 
    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
    all <- rbind(all, tmp)
  }
  all <- all %>% filter(land1_id < land2_id)
  saveRDS(all, file="../data/all-sigs.rds")
} else {
  all <- readRDS("../data/all-sigs.rds")
}

@@

```{r data-profiles}
if (!file.exists("../data/all-profiles.rds")) {
  files <- dir("../data/profiles/", pattern="csv")
  allp <- data.frame()
  for (file in files) {
    tmp <- read.csv(file=file.path(path="../data/profiles", file)) 
    #print(names(tmp))
    if (length(grep("coarse",names(tmp))) == 0)
      tmp$coarse <- gsub(".*coar-(.*).csv", "\\1", file)
    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
    allp <- rbind(allp, tmp)
  }
  allp <- allp %>% filter(land1_id < land2_id)
  allp <- allp %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
  saveRDS(allp, file="../data/all-profiles.rds")
} else {
  allp <- readRDS("../data/all-profiles.rds")
}

@@

if (!file.exists("../data/invprofiles.rds")) {
  files <- dir("../data/inverse-profiles/", pattern="rds")
  allinv <- data.frame()
  for (file in files) {
    tmp <- readRDS(file=file.path(path="../data/inverse-profiles", file)) 
    allinv <- rbind(allinv, tmp)
  }
  allinv <- allinv %>% filter(land1_id < land2_id)
  allinv <- allinv %>% select(-coarseness)
  saveRDS(allinv, file="../data/invprofiles.rds")
} else {
  allinv <- readRDS("../data/invprofiles.rds")
}

@@

```{r data-sig-coars}
if (!file.exists("../data/all-sig-c.rds")) {
  files <- dir("../data/sig_diff_coarseness/", pattern="csv")
  all_sig_c <- data.frame()
  for (file in files) {
    tmp <- read.csv(file=file.path(path="../data/sig_diff_coarseness/", file)) 
    #print(names(tmp))
    if (length(grep("coarse",names(tmp))) == 0)
      tmp$coarse <- gsub(".*coarse-(.*).csv", "\\1", file)
    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
    all_sig_c <- rbind(all_sig_c, tmp)
  }
  all_sig_c <- all_sig_c %>% filter(land1_id < land2_id)
  all_sig_c <- all_sig_c %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
  saveRDS(all_sig_c, file="../data/all-sig-c.rds")
} else {
  all_sig_c <- readRDS("../data/all-sig-c.rds")
}

@@

## Objective and Motivation

- Same Source Matching of Bullet lands
- Evaluate performance of Chumbley Score method when used for Bullet Striations
- Bullet striations have curvature, not present in toolmarks
- Identify Error rates and effect of different parameters on them (In short finding the best error rates possible )

### Structure
- Data being used
- What is the Chumbley Score Method?
- Identifying Best parameter Settings for Bullets
- Modifications to the Algorithm
- Results using \citet{hadler} method and results using Modified method

## Data
-  Ruger P85s Bullet Lands, or Hamby scans (\cite{hamby}) provided by NIST  (85,491 comparisons)
-  Bullet striation marks $\approx$ 2mm | Screwdriver marks $\approx$ 7mm (all chumbley score papers)

### Getting from Profiles to Signatures
\vspace{-0.5em}

- Lowess smoothing or Loess fit residuals $=$ Signatures
- Removes topographic structure (curvature)
- Improve the signal to noise Ratio

\vspace{-1.5em}

```{r sigs-profiles, fig.width = 8, fig.height = 2.5, fig.cap="Bullet land profile (left) and the corresponding signature (right)."}
sigs_graphics<- readRDS("../data/sigs_generate_window.rds")

sigs_graphics.match<- as.data.frame(sigs_graphics[1])
sigs_graphics.nonmatch<- as.data.frame(sigs_graphics[2])
colnames(sigs_graphics.match)<- gsub("^.*\\.","", colnames(sigs_graphics.match))
colnames(sigs_graphics.nonmatch)<- gsub("^.*\\.","", colnames(sigs_graphics.nonmatch))
p1 <- sigs_graphics[[1]] %>% filter(land_id==19) %>%
  ggplot(aes(x = y, y = value)) + geom_line() +
  theme_bw() +
  ylab(expression(paste("Depth (in ",mu,m,")", sep=""))) + 
  xlab(expression(paste("Relative Location (in ",mu,m,")", sep=""))) + 
  ggtitle("Profile")

## Error Rates for toolmarks

\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
  \hline
Research paper & Method & Data Source & False Positives & False Negatives\\
  \hline
 \textbf{\citet{manytoolmarks1}} & Maximized & Screwdrivers & & \\ & Correlation &  & - & - \\ \hline
 \textbf{\citet{chumbley}} & Randomized & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 2.3\% & 8.9\%  \\ \hline
  \textbf{\citet{afte-chumbley}} & Randomized & Slip-joint  &  & \\ & Chumbley Score &  & - & - \\ \hline 
  \textbf{\citet{hadler}} & Deterministic & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 0\% & 6\%  \\ \hline
 \textbf{\citet{manytoolmarks2}} & Similarity Measure & Screwdrivers & & \\ (Different Surfaces-same angle) & Relative Distance Metric & & 5.9\% &  9.4\% \\ (Same Surfaces-same angle)&  & & 0.22\% &  0\% \\
   \hline
\end{tabular}}
\caption{\label{tab:toolmarks-ER} Error Rates for Toolmarks}
\end{table}

## Digitized Striation Marks

- Let $x(t_1)$, $t_1 = 1,2,...T_1$ and $y(t_2)$, $t_2 = 1,2...T_2$ be two digitized marks (where $T_1$ and $T_2$ are not necessarily equal).

- $T_1$ and $T_2$ are the final pixel indexes of each marking. 
Therefore give the respective lengths of the markings.

- Signatures/ Profiles (NIST- Hamby) $\approx$ 1200 pixels (2 mm)
  Screwdriver toolmarks (Chumbley Papers) $\approx$ 9000 pixels (7 mm)
- Markings are smoothed as a preprocessing step
    - Used to remove drift and (sub)class characteristics from individual markings

## Overview

\begin{figure}[hbtp]
\centering
\includegraphics[width=1\textwidth]{imgs/flow-chart-adj-chumbley.png}

\caption{\label{flow-chart} An overview of the chumbley score method}
\end{figure}

## Chumbley Score

- Two steps: Optimization ($1^{st}$) and Validation ($2^{nd}$).
- Windows $\implies$ short segments of the markings
    - Have \textit{predefined sizes}. ($T_1$ or $T_2 >>> w_o >> w_v$)
          1. $w_o$ used in the Optimization step 
          2. $w_v$ used in the Validation step

\vspace{-1em}

#### Optimization step
\vspace{-0.5em}

- Aligns markings horizontally 
- Correlation of all possible windows of size $w_o$ between $x(t_1)$ and $y(t_2)$ computed
- \textit{Possible lag in the 2 markings} \newline Window Pair with maximized correlation $\implies$ \newline Optimal vertical (in-phase) shift of $t_1^o - t_2^o$ 
      - For aligning the two markings. 
\vspace{-1em}
\[
\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x^{w_o} (t_1), y^{w_o}(t_2) \right)
\]

\hspace{-1.5em}
where $t_1^o,t_2^0$ are the respective starting points of $w_o$ in $x(t_1)$ and $y(t_2)$

## 2
- Let $t_1^*$ and $t_2^*$ be relative optimal locations, where $t_i^* = t_i^o/(T_i-w_o)$ for $i=1,2$, such that $t_1^*, t_2^* \in [0,1]$. 
-  Once (sub-)class characteristics are removed, these locations have uniform distribution in $[0,1]$

#### Validations Step
- Two sets of windows of size $w_v$ chosen from both markings (see Figure \ref{fig:win-comparison})

- First set or \textbf{Same Shift}
      - pairs of windows are extracted from the two markings using the optimal vertical shift.
$t_1^o - t_2^o$

- Second set or \textbf{Different Shift}
      - the windows are extracted using a different (out-of-phase) shift.

## In-phase and Out-of-phase

@@

## 2

More precisely, let us define starting points of the windows of validation $s_i^{(k)}$ for each signature $k = 1, 2$ as 
\begin{eqnarray}\label{eq.start}
s^{(k)}_i = 
\begin{cases}
t_k^* + i w_v^* & \text{ for } i < 0 \\
t_k^* + w_ o^* + i w_v^* & \text{ for } i \ge 0,
\end{cases}
\end{eqnarray} for integer values of $i$ with $0 <  s^{(k)}_i \le 1 - w_v^*$
<!-- $0 <  s^{(k)}_i \le T_k - w_v$. -->

## 3

- Same-shift pairs of length $w_v$ are all pairs $$(s_i^{(1)}, s_i^{(2)}) \quad \forall \, i \, \in \, \mathbb{Z}$$ for which both $s_i^{(1)}$ and  $s_i^{(2)}$ are defined. 
- Different-shift pairs are defined as $$(s_i^{(1)}, s_{-i-1}^{(2)}) \quad \forall \, i \, \in \, \mathbb{Z}$$ where both $s_i^{(1)}$ and  $s_{-i-1}^{(2)}$ are defined (see \cref{sketch-same-diff}).

\begin{figure}[hbtp]
\centering
\includegraphics[width=.7\textwidth]{../doc/images/sketch-same.png}

\includegraphics[width=.7\textwidth]{../doc/images/sketch-diff.png}
\caption{\label{sketch-same-diff}Sketch of same-shift pairings  (top) and different-shift pairings (bottom). Filled in rectangles show pairings resulting in correlations, unfilled rectangles are segments without a match.}
\end{figure}

## 4
- No overlaps of pairs within selected marks $\implies$ independence.
- Both same- and different-shift pairs correlations between the  markings are calculated. 
- For Same-Source markings correlations
    - for the in-phase shift should be high
    - for out-of-phase shift should be low.
          - Provide a measure for the base-level correlation to which in-phase shift correlations can be compared.
- The Chumbley score is the Mann Whitney U statistic computed by comparing between in-phase sample and out-of-phase sample.

## Failed Tests

- By definition (equation \autoref{eq.start}), some number of tests fail to produce a result
- Either because the number of eligible same-shift pairs is 0, or the number of different-shift pairs is 0.

\vspace{-0.5em}

###### No Same-shift pairs if
- the optimal locations $t_1^{o}$ and $t_2^{o}$ are in the extremes
    - no segments of size $w_v$ are left on the same sides of the optimal locations, 
    - For all $t_1^{o} < w_v$ and $t_2^{o} > T_2-w_o-w_v$ or \newline $t_1^{o} < T_1- w_o - w_v$ and $t_2^{o} < w_v$

\vspace{-0.5em}

###### No Different Shift Pairs if
- the optimal locations are close to the boundaries
    - \# of different-shift pairs also depends on the placement of the  optimal locations $t_i^o$.
    -  For all $t_i^{(o)} < w_v$ or  $t_i^{(o)} > T_i-w_o- w_v$.

\vspace{-0.5em}

##

#### Failure Rate
\vspace{-1.5em}

\[
\begin{split}
P\left( t_1^{o} < w_v \ \bigcap \ t_2^{o} > T_2-w_o-w_v\right) + \\P\left( t_1^{o} < T_1- w_o - w_v \ \bigcap \ t_2^{o} < w_v\right).
\end{split}
\]

\vspace{-1.5em}

#### Same-shift failure

\begin{figure}[hbtp]
\centering
\includegraphics[width=.7\textwidth]{imgs/same-shift-failure.png}

\caption{\label{same-shift-failure}Sketch of same-shift pairings  (top) when the lag is too large to accomodate a a vaildation window in either of the two signatures}
\end{figure}

## 6

- For different-source matches
    - we can assume $t_1^o$ and $t_2^o$ are independent.
    - Failure rate = $\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}$ 
    - For an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels this probability is about 0.0015.
- For same-source matches, 
    - we expect a strong dependency between optimal locations $t_i^o$
    - $t_1^o$ and $t_2^o$ are not independent of each other
    - A large difference between locations is unlikely
    - Failure Rate = $2 w_v / (T_i-w_o)$, or about 5.6\% for an average length of $T_i$ of 1200 pixels, $w_o = 120$ pixels and $w_v = 30$ pixels.

##  Failure rate for same-source matches
The failure rate for same-source matches can also be written as follows
$$
\begin{split}
P\left( t_1^{o} < w_v \ \bigcup \ t_2^{o} > T_2-w_o-w_v\right) + \\P\left( t_1^{o} < T_1- w_o - w_v \ \bigcup \ t_2^{o} < w_v\right)
\end{split}
$$

Taking first part of the above equation, let $A = t_1^{o} < w_v$ and $B = t_2^{o} > T_2-w_o-w_v$

- Again we assume high correlation between optimal locations
- In order to have at least one same-shift pair or different shift pair
    - \textbf{Either}, if $x \in A$, then $B$ is empty and $P(A \cup B) = P (A)$
    - \textbf{or}, if $x \in B$, then $A$ is empty and $P(A \cup B) = P(B)$

This would hold for the second part too, and would therefore give an expected rate of failure as $2 w_v / (T_i-w_o)$

## Proposed Modification
